Surface correlation functions calculations: improving methodology for discrete images with novel exact continuous approach

### 2023-03-23 paper.tex
Schluss mit den Winkeln!

\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{images/fss-zoomed.png}
  \caption[]{Zoomed image of intersection between the interface $\Gamma$ and its
    translated version $\Gamma'$. $\phi$ is an angle between tangent lines to
    the interfaces. Surface of dashed area is $\frac{\Delta^2}{\sin \phi}$.}
  \label{fig:fss-zoomed}
\end{figure}
Now define a functional sequence $F_{ss}(\bm{x}; \Delta)$:
\begin{align*}
  F_{ss}(\bm{r}; \Delta) &= \int\int f_\Gamma(\bm{x}; \Delta) f_\Gamma(\bm{x}
  + \bm{r}; \Delta) dx dy \\
  &= \sum_{k=1}^N S_k/\Delta^2
\end{align*}
Here $N$ is a number of intersections of $\Gamma$ with itself translated by a
vector $\bm{r}$ and $S_k$ is a surface of each such intersection (see
\cref{fig:Fss-explained}).
When $\Delta$ tends to zero, the boundary at the point of intersection can
be replaced with a tangent line at the point of intersection
(\cref{fig:fss-zoomed}). We get an expression for a surface of
intersection: $S_k = \frac{\Delta^2}{\sin \phi_k}$ where $\phi_k$ is an
acute angle between tangent lines. Hence, we get the following equation:
\begin{equation}
  F_{ss}(\bm{r}) = \lim_{\Delta \to 0} F_{ss}(\bm{r}; \Delta) =
  \sum_{k=1}^N \frac{1}{\sin \phi_k} \label{eq:fss-2d}
\end{equation}
It's easy to show that
$\lim_{\Delta \to 0} f_\Gamma(\bm{x}; \Delta) = |\nabla \chi_A(\bm{x})|$
and therefore \cref{eq:fss-2d} and \cref{eq:fss} define the same function.

@@

\subsection{Surface-surface function for two-dimensional sets with smooth boundary}
\label{sec:cool-algo}
Here we examine two-dimensional sets with smooth boundary described in
\cref{sec:fss-2d} more closely. Suppose a set $S$ is defined as
$S = \left\{ (x, y) \ | \quad f(x, y) \le T \right\}$ where $f(x, y)$ is some function
differentiable at points where $f(x, y) = T$ and $T$ is some threshold. Let's
calculate surface-surface function for this set at a point $(a, b)$. We find all
intersections between a boundary $\Gamma$ and its translated version by solving a
system of equations
\begin{equation*}
  \left\{
  \begin{array}{l}
    f(x, y) = T \\
    f(x-a, y-b) = T    
  \end{array}
  \right.
\end{equation*}
which can be replaced by one equation
\begin{equation}
  (f(x, y) - T)^2 + (f(x-a, y-b) - T)^2 = 0 \label{eq:inter}
\end{equation}
We denote a set of solutions of this equation as $X$. For every $\bm{x}_i \in X$
we find an angle $\phi_i$ between a normal to the boundary and a normal to
the translated boundary as
\begin{equation}
  \cos \phi_i = \frac{(\nabla f(x,y) \vert_{\bm{x}_i}, \nabla f(x-a,y-b)
    \vert_{\bm{x}_i})}{|\nabla f(x,y) \vert_{\bm{x}_i}| |\nabla f(x-a,y-b)
    \vert_{\bm{x}_i}|} \label{eq:angle}
\end{equation}
Here $(\cdot, \cdot)$ denotes a dot product. Gradient of $f$ at the point
$\bm{x_i}$ can be found using automatic differentiation and dual numbers
described in \cref{sec:dual}. Note that acute angle between tangent lines equals
to acute angle between normals, therefore we can find correlation function
$F_{ss}(a, b)$ using the following formula:
\begin{equation}
  F_{ss}(a, b) = \sum_i \frac{1}{\sqrt{1 - \cos^2 \phi_i}} \label{eq:fss-method}
\end{equation}

The algorithm of calculating $F_{ss}(a, b)$ is the following:
\begin{enumerate}
\item Find $X$ which is a set of soulutions of \cref{eq:inter}.
\item For all points in $X$ find cosines between normals to the boundaries using
  \cref{eq:angle}.
\item Calculate $F_{ss}(a, b)$ using \cref{eq:fss-method}.
\end{enumerate}

>>>>

\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{images/fss-zoomed.png}
  \caption[]{Zoomed image of intersection between the interface $\Gamma$ and its
    translated version $\Gamma'$. $\phi$ is an angle between tangent lines to
    the interfaces. Surface of dashed area is
    $\frac{\Delta^2}{\sin \phi}$. Vectors $\bm{a}$ and $\bm{b}$ are unit vectors
    parallel to normals to $\Gamma$ and $\Gamma'$ respectively in the point of
    their intersection.}
  \label{fig:fss-zoomed}
\end{figure}
Now define a functional sequence $F_{ss}(\bm{x}; \Delta)$:
\begin{align*}
  F_{ss}(\bm{r}; \Delta) &= \int\int f_\Gamma(\bm{x}; \Delta) f_\Gamma(\bm{x}
  + \bm{r}; \Delta) dx dy \\
  &= \sum_{k=1}^N S_k/\Delta^2
\end{align*}
Here $N$ is a number of intersections of $\Gamma$ with itself translated by a
vector $\bm{r}$ and $S_k$ is a surface of each such intersection (see
\cref{fig:Fss-explained}).
When $\Delta$ tends to zero, the boundary at the point of intersection can
be replaced with a tangent line at the point of intersection
(\cref{fig:fss-zoomed}). We get an expression for a surface of
intersection: $S_k = \frac{\Delta^2}{\sin \phi_k}$ where $\phi_k$ is an
acute angle between tangent lines. Hence, we get the following equation:
\begin{equation}
  F_{ss}(\bm{r}) = \lim_{\Delta \to 0} F_{ss}(\bm{r}; \Delta) =
  \sum_{k=1}^N \frac{1}{\sin \phi_k} \label{eq:fss-2d-sin}
\end{equation}
It's easy to show that
$\lim_{\Delta \to 0} f_\Gamma(\bm{x}; \Delta) = |\nabla \chi_A(\bm{x})|$
and therefore \cref{eq:fss-2d-sin} and \cref{eq:fss} define the same function.

A less illustrative but more practical way to calculate an expression
$\frac{S_k}{\Delta}$ is the following. Take two vectors $\bm{a}$ and $\bm{b}$
parallel to normals to $\Gamma$ and $\Gamma'$ in $k$-th point of their
intersection such as $|a| = |b| = 1$. If $F_{ss}$ is defined at the point of
consideration then $\bm{a} \ne k\bm{b}$. The pair $(a, b)$ may serve as a basis
with the following transformation from the basis $(0,1), (1,0)$:
\begin{align*}
  x' &= f(x, y) = a_x x + a_y y \\
  y' &= g(x, y) = b_x x + b_y y
\end{align*}
where $\bm{a} = (a_x, a_y)$, $\bm{b} = (b_x, b_y)$. Jacobian determinant for
this transformation is
\begin{equation*}
  J = \left|
  \begin{array}{cc}
    \frac{\partial f}{\partial x} & \frac{\partial f}{\partial y} \\
    \frac{\partial g}{\partial x} & \frac{\partial g}{\partial y} \\
  \end{array}
  \right| =
  \left|
  \begin{array}{cc}
    a_x & a_y \\
    b_x & b_y
  \end{array}
  \right| = a_x b_y - a_y b_x
\end{equation*}

We can rewritte \cref{eq:fss-2d-sin} using $J$ as follows:
\begin{equation}
  F_{ss}(\bm{r}) = \sum_{k=1}^N \frac{1}{|J_k|} \label{eq:fss-2d}
\end{equation}
where $J_k$ is Jacobian determinant computed at $k$-th point of
intersection. This form proves to be much more useful than the version which
contains the angle $\phi$.

@@

\subsection{Surface-surface function for two-dimensional sets with smooth boundary}
\label{sec:cool-algo}
Here we examine two-dimensional sets with smooth boundary described in
\cref{sec:fss-2d} more closely. Suppose a set $S$ is defined as
$S = \left\{ (x, y) \ | \quad f(x, y) \le T \right\}$ where $f(x, y)$ is some function
differentiable at points where $f(x, y) = T$ and $T$ is some threshold. Let's
calculate surface-surface function for this set at a point $(a, b)$. We find all
intersections between a boundary $\Gamma$ and its translated version by solving a
system of equations
\begin{equation*}
  \left\{
  \begin{array}{l}
    f(x, y) = T \\
    f(x-a, y-b) = T    
  \end{array}
  \right.
\end{equation*}
which can be replaced by one equation
\begin{equation}
  (f(x, y) - T)^2 + (f(x-a, y-b) - T)^2 = 0 \label{eq:inter}
\end{equation}
We denote a set of solutions of this equation as $X$. For every $\bm{x}_i \in X$
we find normals to the boundary and the boundary translated by a vector
$(a, b)$. These normals can be found using automatic differentiation and dual
numbers described in \cref{sec:dual}. Then we normalize these normals (i.e. make
their length equal to $1$) and use \cref{eq:fss-2d} to calculate $F_{ss}$.

Summarizing, the algorithm of calculating $F_{ss}(a, b)$ is the following:
\begin{enumerate}
\item Find $X$ which is a set of soulutions of \cref{eq:inter}.
\item For all points in $X$ find normals to the boundaries using automatic
  differentiation.
\item Calculate $F_{ss}(a, b)$ using \cref{eq:fss-2d}.
\end{enumerate}
